Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-05-10 manuscript.tex
edits to intro

  \abstract{The past 15 years have seen increasing experimental investigations of
core pragmatic questions in the ever more active and lively field of
experimental pragmatics. Within experimental pragmatics, many of the
core questions have relied on the operationalization of the theoretical
notion of `implicature rate'. Implicature rate based results have
informed the work on acquisition, online processing, and scalar
diversity, inter alia. Despite its theoretical importance. Implicature
rate has typically been quantified as the proportion of `pragmatic'
judgments in two-alternative forced choice truth value judgment tasks.
Despite its theoretical importance, this linking hypothesis from
implicature rate to behavioral responses has never been extensively
tested. Here we show that two factors dramatically affect the
`implicature rate' inferred from truth value judgment tasks: a) the
number of responses provided to participants; and b) the linking
hypothesis about what constitutes a `pragmatic' judgment. We argue that
it is time for the field of experimental pragmatics to engage more
seriously with its foundational assumptions about how theoretical
notions map onto behaviorally measurable quantities, and present a
sketch of an alternative linking hypothesis that derives behavior in
truth value judgment tasks from probabilistic utterance expectations.}
  \keywords{scalar implicature; methodology; linking hypothesis; experimental
pragmatics; truth value judgment task \\

@@

\emph{Scale:} \(<\)all, some\(>\)

@@

\emph{Scale:} \(<\)and, or\(>\)

A listener, upon observing the utterances in (1-2a) typically infers
that the speaker intended to convey the meanings in (1-2b),
respectively. Since Grice (1975), the agreed-upon abstract
rationalization the listener could give for their inference goes
something like this: the speaker could have made a more informative
statement by producing the stronger alternative (e.g., \emph{All of her
pets are cats.} in (1)). If the stronger alternative is true, they
should have produced it to comply with the Cooperative Principle. They
chose not to. I believe the speaker knows whether the stronger
alternative is true. Hence, it must not be true. The derivation
procedure for ad hoc exhaustive inferences such as in (3) is assumed to
be calculable in the same way as for scalar implicatures, though the
scale is assumed to be contextually driven.

@@

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as `Some, and possibly
all, elephants are mammals' or \enquote{You got some, and possibly all,
of the gumballs}.

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of FALSE responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute 1, 2, or 3 strawberries to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average fewer strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options -- TRUE and FALSE -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly FALSE responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypothesiss}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues. First,
following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypothesiss.
Next, we discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypothesiss in a rigorous and motivated way.

>>>>

  \abstract{The past 15 years have seen increasing experimental investigations of
core pragmatic questions in the ever more active and lively field of
experimental pragmatics. Within experimental pragmatics, many of the
core questions have relied on the operationalization of the theoretical
notion of `implicature rate'. Implicature rate based results have
informed the work on acquisition, online processing, and scalar
diversity, inter alia.  Implicature
rate has typically been quantified as the proportion of `pragmatic'
judgments in two-alternative forced choice truth value judgment tasks.
Despite its theoretical importance, this linking hypothesis from
implicature rate to behavioral responses has never been extensively
tested. Here we show that two factors dramatically affect the
`implicature rate' inferred from truth value judgment tasks: a) the
number of responses provided to participants; and b) the linking
hypothesis about what constitutes a `pragmatic' judgment. We argue that
it is time for the field of experimental pragmatics to engage more
seriously with its foundational assumptions about how theoretical
notions map onto behaviorally measurable quantities, and present a
sketch of an alternative linking hypothesis that derives behavior in
truth value judgment tasks from probabilistic utterance expectations.}
  \keywords{scalar implicature; methodology; linking hypothesis; experimental
pragmatics; truth value judgment task \\

@@

\emph{Scale:} $\langle$all, some$\rangle$

@@

\emph{Scale:} $\langle$and, or$\rangle$

A listener, upon observing the utterances in (1-2) typically infers
that the speaker intended to convey the meanings listed as \emph{Implicature}s,
respectively. Since Grice (1975), the agreed-upon abstract
rationalization the listener could give for their inference goes
something like this: the speaker could have made a more informative
statement by producing the stronger alternative (e.g., \emph{All of her
pets are cats} in (1)). If the stronger alternative is true, they
should have produced it to comply with the Cooperative Principle. They
chose not to. Assuming the speaker knows whether the stronger
alternative is true, it must not be true. The derivation
procedure for ad hoc exhaustivity inferences such as in (3) is assumed to
be calculable in the same way as for scalar implicatures, though the
scale is assumed to be contextually driven.

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  She owns a cat.
\end{enumerate}

\emph{Implicature:} She owns only a cat.

\emph{Scale:} $\langle$cat and dog, cat$\rangle$

@@

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as ``Some, and possibly
all, elephants are mammals'' or \enquote{You got some, and possibly all,
of the gumballs}.

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute 1, 2, or 3 strawberries to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average fewer strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options -- ``TRUE'' and ``FALSE' -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly ``FALSE' responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypotheses}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

### 2018-05-10 writeup_jd.Rmd
edits to intro

  - name          : "Masoud Jasbi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Margaret Jacks Hall, Building 460 Rm. 127, Stanford, CA, 94305-2150"
    email         : "masoudj@stanford.edu"
  - name          : "Brandon Waldon"
    affiliation   : "1"
  - name          : "Judith Degen"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Stanford University, Department of Linguistics"
author_note: |
  The authors declare that they have no affiliations with, or involvement in any organization or entity with any financial interest, or non-financial interest in the subject matter or materials discussed in this manuscript. 
abstract: |
  The past 15 years have seen increasing experimental investigations of core pragmatic questions in the ever more active and lively field of experimental pragmatics. Within experimental pragmatics, many of the core questions have relied on the operationalization of the theoretical notion of 'implicature rate'. Implicature rate based results have informed the work on acquisition, online processing, and scalar diversity, inter alia. Despite its theoretical importance. Implicature rate has  typically been quantified as the proportion of 'pragmatic' judgments in two-alternative forced choice truth value judgment tasks. Despite its theoretical importance, this linking hypothesis from implicature rate to behavioral responses has never been extensively tested. Here we show that two factors dramatically affect the 'implicature rate' inferred from truth value judgment tasks: a) the number of responses provided to participants; and b) the linking hypothesis about what constitutes a 'pragmatic' judgment. We argue that it is time for the field of experimental pragmatics to engage more seriously with its foundational assumptions about how theoretical notions map onto behaviorally measurable quantities, and present a sketch of an alternative linking hypothesis that derives behavior in truth value judgment tasks from probabilistic utterance expectations.
keywords          : "scalar implicature; methodology; linking hypothesis; experimental pragmatics; truth value judgment task"
wordcount         : "9037"
bibliography      : ["r-references.bib"]
figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: inline
---

@@

(1) Some of her pets are cats.

*Implicature:* Some, but not all, of her pets are cats.

*Scale:* $<$all, some$>$

(2) She owns a cat or a dog.

*Implicature:* She owns a cat or a dog, but not both.

*Scale:* $<$and, or$>$

@@

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative or weakly informative priors [^3]. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^5].

@@

[^5]: You can access our pre-registration at [https://aspredicted.org/tq3sz.pdf](https://aspredicted.org/tq3sz.pdf)

@@

$P_{S_1}(u | s) \propto exp(\alpha(log(L_0(s | u)) - C(u)))$

@@

```{r speakerprobs, message=FALSE, warning=FALSE, fig.cap = "Speaker probabilities of utterances on the exhaustive and scalar trials, as obtained using the model described in this section."}

@@

speaker_probs %>%
  ggplot(aes(x=u, y=prob)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_wrap(~s) +
  labs(x="", y= paste("Speaker probability ", "S(u | s)"))+
  # theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE)
```

>>>>

  - name          : "Masoud Jasbi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "XXX Postal address"
    email         : "XXX my@email.com"
  - name          : "Brandon Waldon"
    affiliation   : "1"
  - name          : "Judith Degen"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Stanford University"
author_note: |
  XXX Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  The past 15 years have seen increasing experimental investigations of core pragmatic questions in the ever more active and lively field of experimental pragmatics. Within experimental pragmatics, many of the core questions have relied on the operationalization of the theoretical notion of 'implicature rate'. Implicature rate based results have informed the work on acquisition, online processing, and scalar diversity, inter alia. Despite its theoretical importance. Implicature rate has  typically been quantified as the proportion of 'pragmatic' judgments in two-alternative forced choice truth value judgment tasks. Despite its theoretical importance, this linking hypothesis from implicature rate to behavioral responses has never been extensively tested. Here we show that two factors dramatically affect the 'implicature rate' inferred from truth value judgment tasks: a) the number of responses provided to participants; and b) the linking hypothesis about what constitutes a 'pragmatic' judgment. We argue that it is time for the field of experimental pragmatics to engage more seriously with its foundational assumptions about how theoretical notions map onto behaviorally measurable quantities, and present a sketch of an alternative linking hypothesis that derives behavior in truth value judgment tasks from probabilistic utterance expectations.
keywords          : "scalar implicature; methodology; linking hypothesis; experimental pragmatics; truth value judgment task"
wordcount         : "X"
bibliography      : ["r-references.bib"]
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: inline
---

@@

1. 
  + *Utterance:* Some of her pets are cats.
  + *Implicature:* Some, but not all, of her pets are cats.
  + *Scale:* <all, some>

2. 
  + *Utterance:* She owns a cat or a dog.
  + *Implicature:* She owns a cat or a dog, but not both.
  + *Scale:* <and, or>

@@

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative or weakly informative priors [^3]. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

@@

[^3]: You can access our pre-registration at [https://aspredicted.org/tq3sz.pdf](https://aspredicted.org/tq3sz.pdf)

@@

    $P_{S_1}(u | s) \propto exp(\alpha(log(L_0(s | u)) - C(u))) $

@@

```{r speakerprobs, message=FALSE, warning=FALSE,fig.cap = "Speaker probabilities of utterances on the exhaustive and scalar trials, as obtained using the model described in this section."}

@@

speaker_probs %>%
  ggplot(aes(x=u, y=prob)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_wrap(~s) +
  labs(x="", y= paste("Speaker probability ", "S(u | s)"))+
  # theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1))
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE)
```
